%\input{template.tex}
%\begin{document}

First, I would like to express my special appreciation and gratitude to my advisor Dr. Haimeng Zhang for his motivation and willingness to share the knowledge he aquired over the years. His door was always opened to me for any questions; without his support and guidance this dissertation would not have been possible. Dr. Zhang's advise was priceless, it helped me not only to grow as a young researcher but also as a teacher, a professional and a leader. 

I would like to thank my dissertation committee members, Dr. Sat Gupta, Dr. Scott Richter, Dr. Xiaoli Gao, and Dr. Shan Suthaharan, for been in my committee, sharing their ideas, suggestions and comments to improve this dissertation. Especially, I would  like to thank Mathematics and Statistics Department for providing me a Teaching Assitantship to support my studies. Further, I would like to thank Dr. Scott Richter for engaging me in his consulting projects and for providing a Research Assiatantship during summer through Statistical Consulting Center. I wish to thank all faculty, staff and my friends at Mathematics and Statistics Department for all their support and fun time at UNCG.  

A special thanks to my mother, sister, relatives and friends for their endless support. My father is not with me today to share my results but I would like to dedicate this dissertation to him for his inspiration and encouragement. Finally, I would like to express appreciation to my beloved wife Lakma Fernando who supported me every possible way in this journey and my two kids Ethmi and Kethmin who keep me busy and for bringing happiness and joy to our family.    



%\end{document}

% , without the Assistantship my studies wouldn't have been possible.